---
title: "Homework 4"
author: "Mason Yates"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
---

## Instructions
Answer the following questions and/or complete the exercises in RMarkdown. Please embed all of your code and push the final work to your repository. Your report should be organized, clean, and run free from errors. Remember, you must remove the `#` for any included code chunks to run. Any plots must have appropriate titles and aesthetics.  

## Load the tidyverse
```{r}
library("tidyverse")
options(scipen = 999) #disable scientific notation
```

For this homework, we will be using the `msleep` dataset from the `ggplot2` package. This dataset contains information about the sleep habits of various mammals. [Reference](https://ggplot2.tidyverse.org/reference/msleep.html): V. M. Savage and G. B. West. A quantitative, theoretical framework for understanding mammalian sleep. Proceedings of the National Academy of Sciences, 104 (3):1051-1056, 2007.

**1. Use a summary function of your choice to get an idea of the structure of the msleep data.**
```{r}
msleep <- msleep
glimpse(msleep)
```

**2. What are the names of the variables? To get a better idea of what the variables mean, you can use `?msleep`.**

The names of the variables are:
order,
conservation,
sleep_total,
sleep_rem,
sleep_cycle,
awake,
brainwt,
bodywt

**3. Make a plot that shows the number of mammals in each vore type. Which vore type is most represented in the data?**

Herbivores are the most represented.

```{r}
ggplot(data=msleep,
       mapping=aes(x=vore)
       )+
  geom_bar(mapping=aes(fill=vore))+
  scale_x_discrete(na.translate=F)+
  labs(title="Mammals of Each Vore",
       x="Vore",
       y="Mammal Count")
```


**4. What is the average length of time that mammals sleep?**

```{r}
average_sleep <- mean(msleep$sleep_total, na.rm=T)
average_rounded <- signif(average_sleep, digits=4)
average_rounded
```

The average length of sleep is 10.43 hours

**5. Use `summary()` to get a quick idea of the range of values for the variable `bodywt`.**

```{r}
summary(msleep$bodywt)
```


**6. Build a plot that shows the range of `bodywt.` by vore type. Because the range of `bodywt` is so variable, you will need to scale the y-axis. You do this by adding `scale_y_log10()` as another layer.**

```{r}
ggplot(data=msleep,
       mapping=aes(x=vore, y=bodywt)
       )+
  geom_boxplot(mapping=aes(fill=vore))+
  scale_y_log10()+
  scale_x_discrete(na.translate=F)+
  labs(title="Body Weights (kg) of Vore Types",
       x="Vore type",
       y="Body Weight (kg)")
```


**7. Is there a relationship between body weight and total sleep time? Add appropriate labels and a title to your plot.**

Yes, there is a relationship between body weight and total sleep time. Higher body weights indicate lower total sleep time.

```{r}
ggplot(data=msleep,
       mapping=aes(x=bodywt, y=sleep_total)
       )+
  geom_point(mapping=aes(color=vore))+
  geom_smooth(method=lm, color="black")+
  scale_x_log10()+
  labs(title="Body Weight (kg) vs Total Sleep (Hrs)",
       x = "Body Weight (kg)",
       y = "Total Sleep (Hrs)")
```


**8. Which vore type tends to sleep the most? Build a plot to support your answer.**

Insectivores tend to sleep the most.

```{r}
ggplot(data=msleep,
       mapping=aes(x=vore, y=sleep_total)
       )+
  geom_boxplot(mapping=aes(fill=vore))+
  scale_x_discrete(na.translate=F)+
  labs(title="Sleep Times (Hrs) of Vore Type",
       x="Vore Type",
       y="Sleep Time (Hrs)")
```


**9. Is there a relationship between brain weight and body weight? Build a plot to support your answer.**

Yes, there is a relationship between brain weight and body weight. As body weight increases, brain weight is likely to increase.

```{r}
ggplot(data=msleep,
       mapping=aes(y=brainwt, x=bodywt)
       )+
  geom_point(mapping=aes(color=vore))+
  geom_smooth(method=lm, color="black")+
  scale_x_log10()+
  labs(title="Body Weight (kg) vs Brain Weight (kg)",
       x="Body Weight (kg)",
       y="Brain Weight (kg)")
  
```


**10. Build one plot of your choice that provides additional insight into the msleep dataset. Be sure to include appropriate labels and a title.**

```{r}
brown_green_palette <- c(
  "#2F5D50",  
  "#6B8E6E",
  "#8A9A5B",
  "#7A5C47",
  "#5B3A29", 
  "#C8B7A6" 
)
conservation_labels <- c(
  "lc" = "Least Concern",
  "nt" = "Near Threatened",
  "vu" = "Vulnerable",
  "en" = "Endangered",
  "domesticated" = "Domesticated",
  "cd" = "Conservation 
Dependent"
)

ggplot(data=msleep,
       mapping=aes(x=conservation, fill=conservation)
       )+
  geom_bar()+
  scale_fill_manual(values=brown_green_palette)+ #it looks good now!
  scale_x_discrete(labels=conservation_labels, na.translate=F)+ #it has labels!
  labs(title="Count of Present Conservation Statuses",
       x="Conservation Status",
       y="Count")
```


## Knit and Upload
Please knit your work as an .html file and upload to Canvas. Homework is due before the start of the next lab. No late work is accepted. Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  