---
title: "Homework 10"
author: "Mason Yates"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: spacelab
    toc: no
---

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!    

## Background
In the `data` folder, you will find data about shark incidents in California between 1950-2022. The [data](https://catalog.data.gov/dataset/shark-incident-database-california-56167) are from: State of California- Shark Incident Database.   

## Load the libraries
```{r message=FALSE, warning=FALSE}
library("tidyverse")
library("janitor")
```

## Load the data
Run the following code chunk to import the data. In this chunk, I am removing NA's.
```{r message=FALSE, warning=FALSE}
sharks <- 
  read_csv("data/SharkIncidents_1950_2022_220302.csv", 
           na=c("NA", "N/A", "Unknown", "NOT COUNTED", "", " ")) %>% 
  clean_names()
```

```{r}
sharks <- sharks %>% 
  mutate(across(c(month, day, year, time), as.factor))
```

## Questions
1. Are there any "hotspots" for shark incidents in California? Show the top five counties with the most incidents.
```{r}
sharks %>% 
  count(
    county,
    sort = TRUE
  ) %>% 
  slice_max(n, n = 5)
```

2.  Make a plot that shows the total number of incidents per county.
```{r}
county_palette <- c(
  "San Diego" = "#2F5D50",
  "Monterey" = "#3E6B5C",
  "San Luis Obispo" = "#4A7A67",
  "San Francisco" = "#5A8F76",
  "Sonoma" = "#6B9B7C",
  "Los Angeles" = "#7AA58B",
  "Marin" = "#8AA892",
  "Santa Cruz" = "#9CB49B",
  "Island - Farallones" = "#AFC6A3",
  "San Mateo" = "#C0D3B2",
  "Santa Barbara" = "#7A5C47",
  "Mendocino" = "#8B6452",
  "Humboldt" = "#9C735D",
  "Island - San Miguel" = "#AD8268",
  "Orange" = "#BD9174",
  "Del Norte" = "#CEA181",
  "Island - Santa Barbara" = "#DDB18F",
  "Island - San Nicolas" = "#EEC1A0",
  "Ventura" = "#5B4A3A",
  "Island - Catalina" = "#6E7D6A",
  "Island - Santa Cruz" = "#8A9A5B",
  "Island - Santa Rosa" = "#B39B6B"
)

sharks %>% 
  ggplot(
    mapping = aes(
      y = county,
      fill = county
    )
  )+
  geom_bar()+
  scale_fill_manual(
    values = county_palette
  )+
  labs(
    title = "Shark Attacks per County",
    x = "Number of Shark Attacks",
    y = "County",
  )+
    theme(legend.position = "none")
```

3. Are there months of the year when incidents are more likely to occur? Make a plot that shows the total number of incidents by month. Which month has the highest number of incidents?  

Shark incidents are more likely to occur between August and October, with the month of October having the most recorded incidents between 1950 and 2022.

```{r}
month_palette <- c(
  "1"  = "#2F5D50",
  "2"  = "#3E6B5C",
  "3"  = "#4A7A67",
  "4"  = "#5A8F76",
  "5"  = "#6B9B7C",
  "6"  = "#7AA58B",
  "7"  = "#8AA892",
  "8"  = "#9CB49B",
  "9"  = "#8C6A4A",
  "10" = "#9C735D",
  "11" = "#AD8268",
  "12" = "#5B4A3A"
)

sharks %>% 
  ggplot(
    mapping = aes(
      y = month,
      fill = month
    )
  )+
  geom_bar()+
  scale_y_discrete(
    labels = month.name
  )+
  scale_fill_manual(
    values = month_palette
  )+
  labs(
    title = "Total Shark Incidents each Month from 1950 to 2022",
    x = "Count",
    y = "Month"
  )+
  theme(legend.position = 'none')
```

```{r}
```

4. Is there a time of day when most attacks occur? Make a plot that shows the total number of incidents by time of day. Which time of day has the highest number of incidents?

Most attacks occur at 11:00. 

```{r}
sharks %>%
  filter(!time %in% c("afternoon", "early am"), !is.na(time)) %>%
  ggplot(
    mapping = aes(
      x = time,
    )
  )+
  geom_bar(fill = "orange")+  
  theme(
    axis.text.x = element_text(
                                angle = 90, 
                                hjust = 0.85)
    )+
  labs(
    title = "Time of Shark Incidents from 1950 to 2022",
    x = "Time",
    y = "Count"
  )
```

5. How do the number and types of injuries compare by county? Make a table that shows the number of injury types by county. Which county has the highest number incidents? (hint: your table should be in wide format)

San Diego county has the highest number of incidents.

```{r}
shark_wide <- sharks %>%
  select(county, injury) %>%
  mutate(injury = str_remove(injury, "\\*")) %>% 
  group_by(county, injury) %>% 
  summarize(
    injury_count = n(),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = injury,
    values_from = injury_count
  ) %>%
  group_by(county) %>% 
  mutate(
    total.incidents = sum(c(minor,major,fatal,none), na.rm = TRUE)
  ) %>% 
  relocate("minor", "fatal", "none", .after = "major") %>% 
  arrange(desc(total.incidents))
shark_wide
```

6. Use the table from #5 to make a plot that shows the total number of incidents by county.
```{r}
shark_wide %>% 
  ggplot(
    mapping = aes(
      y = county,
      x = total.incidents,
      fill = county
    )
  )+
  geom_col()+
  scale_fill_manual(
    values = county_palette
  )+
  labs(
    title = "Total Shark Incidents per County from 1950 to 2022",
    x = "County",
    y = "Total Incidents"
  )+
  theme(
    legend.position = "none"
  )
```

7. In the data, `mode` refers to a type of activity. Which activity is associated with the highest number of incidents?

Surfing / Boarding is the activity with the highest number of incidents.

```{r}
sharks %>% 
  count(mode, sort = TRUE)
```

8. Make a plot that compares the number of incidents by activity.
```{r}
sharks %>% 
  ggplot(
    mapping = aes(
      x = mode)
  )+
  geom_bar(fill = "orange")+
  theme(
    axis.text.x = element_text(
                                angle = 30, 
                                hjust = 0.95)
    )+
  labs(
    title = "Shark Incidents per Activity Type from 1950 to 2022",
    x = "Activity Type",
    y = "Count"
  )
```

9. Which shark species is involved in the highest number of incidents?  

Great White sharks are involved in the highest number of incidents.

```{r}
sharks %>% 
  count(species, sort = TRUE)
```

10. Are all incidents involving Great White's fatal? Make a plot that shows the number and types of incidents for Great White's only.  
```{r}
sharks %>% 
  filter(species == "White") %>%
  mutate(injury = str_remove(injury, "\\*")) %>% 
  ggplot(
    mapping = aes(
      x = injury
    )
  )+
  geom_bar(fill = "orange")+
  labs(
    title = "Injury Types of Great White Incidents between 1950 and 2022",
    x = "Injury Type",
    y = "Number of Incidents"
  )
```

## Knit and Upload
Please knit your work as an .html file and upload to Canvas. Homework is due before the start of the next lab. No late work is accepted. Make sure to use the formatting conventions of RMarkdown to make your report neat and clean! 